---
title: "Discovering a Global Data Model from Event Logs - Complementary Notebook"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(here)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(igraph)
library(textshape)
library(RVenn)
```


```{r functions, echo=FALSE}
a2a <- function(df, relative = FALSE) {
  if(relative == FALSE){
    df %>% 
      group_by(Activity) %>%
      summarise(across(everything(), ~ sum(!is.na(.x))))
  } else {
    df %>%
      group_by(Activity) %>%
      summarise(across(everything(), ~ sum(!is.na(.x)) / n()))
  }
}

dropSuperfluousAttributes <- function(df, dropCols = NULL) {
  if(is.null(dropCols)) {
    dropCols <- c("Case.ID", "Resource", "Timestamp", "Variant", "Lifecycle")
  }
  df %>% select(-contains(dropCols))
}

replaceEmptyWithNA <- function(df, keepZeroes = FALSE) {
  df[df == ""] <- NA
  if(keepZeroes == FALSE) df[df == "0"] <- NA
  df
}

MLAD.Deep.setup <- function(a2a) {
  df <- slice(a2a, 0)
  df <- bind_rows(df, a2a %>% select(Activity))
  df[is.na(df)] <- 0
  df
}

MLAD.Deep.fill <- function(scaffold, a2a) {
  columns <- colnames(a2a)
  rows <- rownames(a2a)
  for(c in columns) {
    for(r in rows) {
      scaffold[r, c] <- a2a[r, c]
    }
  }
  scaffold
}

MLAD <- function(a, b) {
  sort(intersect(colnames(a), colnames(b)))
}
```

This is the complementary notebook for the report "Discovering Global Data Models from Event Logs".


## Loading of Datasets

For reproducibility, this section shows the datasets' preprocessing.
 
### Offer Applications (BPI 2017)

Meta-attributes are removed, empty strings are coerced to NA. Note, that this dataset makes a distinction between 0 and empty strings, so zeroes are not coerced to NA as they are a value.

```{r dataLoading Offers}

OfferApplications <- read.csv(here("data", "bpi17.csv"))
OfferApplications <- dropSuperfluousAttributes(OfferApplications, c("Resource", "Timestamp", "Variant", "Lifecycle", "Event", "Action"))
OfferApplications <- replaceEmptyWithNA(OfferApplications, keepZeroes = TRUE)
Offers <- read.csv(here("data", "bpi17_offer_log.csv"))
Offers <- dropSuperfluousAttributes(Offers, c("Resource", "Timestamp", "Variant", "Lifecycle", "Event", "Action"))
Offers <- replaceEmptyWithNA(Offers, keepZeroes = TRUE)
```

### Permits (BPI 2020)

Meta-attributes are removed, empty strings and zeroes are coerced to NA.

```{r dataLoading Permits}
Permits <- read.csv(here("data", "PermitLog.csv"))
Permits <- replaceEmptyWithNA(Permits)
Permits <- dropSuperfluousAttributes(Permits)

InternationalDeclarations <- read.csv(here("data", "InternationalDeclarations.csv"))
InternationalDeclarations <- replaceEmptyWithNA(InternationalDeclarations)
InternationalDeclarations <- dropSuperfluousAttributes(InternationalDeclarations)

DomesticDeclarations <- read.csv(here("data", "DomesticDeclarations.csv"))
DomesticDeclarations <- replaceEmptyWithNA(DomesticDeclarations)
DomesticDeclarations <- dropSuperfluousAttributes(DomesticDeclarations)

RequestForPayment <- read.csv(here("data", "RequestForPayment.csv"))
RequestForPayment <- replaceEmptyWithNA(RequestForPayment)
RequestForPayment <- dropSuperfluousAttributes(RequestForPayment)

PrepaidTravelCost <- read.csv(here("data", "PrepaidTravelCost.csv"))
PrepaidTravelCost <- replaceEmptyWithNA(PrepaidTravelCost)
PrepaidTravelCost <- dropSuperfluousAttributes(PrepaidTravelCost)

```


```{r dataLoading Mock, echo = FALSE}
#eventlogmockP1 <- read.csv("D:/uni/bpm/eventlogmockP1.csv")
#eventlogmockP1[eventlogmockP1 == ""] <- NA
#eventlogmockP2 <- read.csv("D:/uni/bpm/eventlogmockP2.csv")
#eventlogmockP2[eventlogmockP2 == ""] <- NA
```

## Algorithmic steps to generate a global data model

### Offer Applications (BPI 2017)

Event Logs are translated to A2A Matrices. MLAD Matrices are combined by union.

```{r bpi2017}
OfferApplications.a2a <- a2a(OfferApplications)
OfferApplications.a2a.relative <- a2a(OfferApplications, relative = TRUE)
Offers.a2a <- a2a(Offers)


cat(c("Class attributes found by MLAD:", "\n", MLAD(Offers, OfferApplications), "\n", "\n"))

Offer.union <- bind_rows(OfferApplications, Offers)
Offer.union.a2a <- a2a(Offer.union)

Offer.union.a2a.remaining <- Offer.union.a2a[ , !(colnames(Offer.union.a2a) %in% MLAD(Offers, OfferApplications))]
cat(c("Remaining attributes found by FCSE:", "\n", colnames(Offer.union.a2a.remaining), "\n"))

```
### Permits (BPI 2020)

Event logs are translated to A2A Matrices.

```{r bpi2020}
Permits.a2a <- a2a(Permits)
DomesticDeclarations.a2a <- a2a(DomesticDeclarations)
InternationalDeclarations.a2a <- a2a(InternationalDeclarations)
PrepaidTravelCost.a2a <- a2a(PrepaidTravelCost)
RequestForPayment.a2a <- a2a(RequestForPayment)

permits.union <- bind_rows(DomesticDeclarations, InternationalDeclarations, Permits, PrepaidTravelCost, RequestForPayment)
permits.union.a2a <- a2a(permits.union)
```


### Multi-log Access Detection for the BPI 2020 Challenge

MLAD is performed pair-wise on all logs. 
```{r bpi2020 mlad}
cat(c("Class candidates:", "\n",
  MLAD(Permits, InternationalDeclarations), "\n", 
  MLAD(Permits, DomesticDeclarations), "\n",
  MLAD(Permits, RequestForPayment), "\n",
  MLAD(Permits, PrepaidTravelCost), "\n",
  
  MLAD(InternationalDeclarations, DomesticDeclarations), "\n",
  MLAD(InternationalDeclarations, RequestForPayment), "\n",
  MLAD(InternationalDeclarations, PrepaidTravelCost), "\n",
  
  MLAD(DomesticDeclarations, RequestForPayment), "\n",
  MLAD(DomesticDeclarations, PrepaidTravelCost), "\n",
  
  MLAD(RequestForPayment, PrepaidTravelCost), "\n"
  )
)

removeAttributes <- list(MLAD(Permits, InternationalDeclarations),
                    MLAD(Permits, DomesticDeclarations),
                    MLAD(Permits, RequestForPayment),
                    MLAD(Permits, PrepaidTravelCost),
                    MLAD(InternationalDeclarations, DomesticDeclarations),
                    MLAD(InternationalDeclarations, RequestForPayment),
                    MLAD(InternationalDeclarations, PrepaidTravelCost),
                    MLAD(DomesticDeclarations, RequestForPayment),
                    MLAD(DomesticDeclarations, PrepaidTravelCost),
                    MLAD(RequestForPayment, PrepaidTravelCost))
ra <- Venn(removeAttributes)
ra <- unite(ra)
removeAttributes <- strsplit(ra, " ")
permits.union.a2a <- permits.union.a2a[ , !(colnames(permits.union.a2a) %in% removeAttributes)]
## Still 167 attributes left after MLAD. 
## Remove Multi-Column Attributes.

permits.union.a2a <- permits.union.a2a[ , 1:14]
cat(c("Attributes found by FCSE:", "\n",
      sort(colnames(permits.union.a2a))))
```

### Multi-log Access Detection in the Reduced BPI 2020 Dataset. 

Since results on the full Permits dataset are convoluted, the attributes are reduced.
Reductions are based on "Multi-perspective Analysis of Approval Processes based on Multiple Event Logs", by Bano et. al:
Many rarely accessed attributes are removed, as well as multi-column attributes of the type "attributeName_0", ... "attributesName_n".

```{r reduced bpi2020 mlad}
reducedPermits <- Permits %>% select(Activity, BudgetNumber, Overspent, OverspentAmount, ProjectNumber, RequestedBudget, TotalDeclared, id)
reducedDomesticDeclarations <- DomesticDeclarations %>% select(Activity, Amount, BudgetNumber, id)
reducedInternationalDeclarations <- InternationalDeclarations %>% select(Activity, Amount, BudgetNumber, id)
reducedRequestForPayment <- RequestForPayment %>% select(Activity, Project, RequestedAmount, Rfp_id)
reducedPrepaidTravelCost <- PrepaidTravelCost %>% select(Activity, Project, RequestedAmount, Rfp_id)


reducedPermits.a2a <- a2a(reducedPermits)
reducedDomesticDeclarations.a2a <- a2a(reducedDomesticDeclarations)
reducedInternationalDeclarations.a2a <- a2a(reducedInternationalDeclarations)
reducedRequestForPayment.a2a <- a2a(reducedRequestForPayment)
reducedPrepaidTravelCost.a2a <- a2a(reducedPrepaidTravelCost)

cat(c("Class candidates:", "\n",
  MLAD(reducedPermits, reducedInternationalDeclarations), "\n",
  MLAD(reducedPermits, reducedDomesticDeclarations), "\n",
  MLAD(reducedPermits, reducedRequestForPayment), "\n",
  MLAD(reducedPermits, reducedPrepaidTravelCost), "\n",
  
  MLAD(reducedInternationalDeclarations, reducedDomesticDeclarations), "\n",
  MLAD(reducedInternationalDeclarations, reducedRequestForPayment), "\n",
  MLAD(reducedInternationalDeclarations, reducedPrepaidTravelCost), "\n",
  
  MLAD(reducedDomesticDeclarations, reducedRequestForPayment), "\n",
  MLAD(reducedDomesticDeclarations, reducedPrepaidTravelCost), "\n",
  
  MLAD(reducedRequestForPayment, reducedPrepaidTravelCost), "\n"
))
```

Two sets of attributes are found that are not fully subsets of another set:

```{r reduced bpi2020 results}



reducedRfp.union <- bind_rows(reducedRequestForPayment, reducedPrepaidTravelCost)
reducedRfp.union <- reducedRfp.union[ , !(colnames(reducedRfp.union) %in% MLAD(reducedRequestForPayment, reducedPrepaidTravelCost))]

reducedIntDom.union <- bind_rows(reducedInternationalDeclarations, reducedDomesticDeclarations)
reducedIntDom.union <- reducedIntDom.union[ , !(colnames(reducedIntDom.union) %in% MLAD(reducedInternationalDeclarations, reducedDomesticDeclarations))] 

# All attributes of the pairs were shared, so after removing the attributes the datasets are empty;
# Only reducedPermits remains.

cat(c("First found class:", "\n",
    MLAD(reducedInternationalDeclarations, reducedDomesticDeclarations), "\n", "\n",
    "Second class:", "\n",
    MLAD(reducedRequestForPayment, reducedPrepaidTravelCost), "\n", "\n",
    "Remaining class by FCSE:", "\n",
    colnames(reducedPermits.a2a), "\n"))


```

```{r, include = FALSE, eval = FALSE}
names.1 <- colnames(EL.1)
names.2 <- colnames(EL.2)


cat("Columns only in log.1:", names.1[!names.1 %in% names.2], "\n",
    "Columns only in log.2:", names.2[!names.2 %in% names.1])
```

```{r Approach 0, include=FALSE, eval = FALSE}
#Rule 1 not applicable

#Rule 3
EL.offerless <- EL.1[!EL.1$Activity == "O_Create Offer", ]
# only "O_Create Offer" has any values in FirstWithdrawalAmount, NumberOfTerms, Accepted, MonthlyCost, Selected, CreditScore, OfferedAmount

EL.NAfree <- EL.offerless[!is.na(EL.offerless$CreditScore), ] #empty

# EL COMPLEMENT E_A1 => E_A2 = NA 
# => E_A1 UNION E_A2 = E_A1
# => "O_Create Offer" is an isolated Activity with non-isolated attributes (FirstWithdrawalAmount, NumberOfTerms, Accepted, MonthlyCost, Selected, CreditScore, OfferedAmount), with ONE class.


#-------------------------------------------
#Rule 2
EL.OfferID.exists <- EL.1[!is.na(EL.1$OfferID), ]
#head(EL.OfferID.exists)
activities.offerID.access <- unique(EL.OfferID.exists$Activity)
#cat(c("Rule 2 fulfilled by: ", activities.offerID.access, "\n")) #Activities accessing Attr "OfferID"


#Rule 4
EL.Remaining <- EL.offerless[! EL.offerless$Activity %in% activities.offerID.access, ]
#cat(c("OfferID empty?", sum(is.na(EL.Remaining$OfferID)) == nrow(EL.Remaining), "\n"))
#cat(c("OfferedAmount empty?", sum(is.na(EL.Remaining$OfferedAmount)) == nrow(EL.Remaining)))
# => Everything remaining accesses everything



#===>
#3. FirstWithdrawalAmount, NumberOfTerms, Accepted, MonthlyCost, Selected, CreditScore, OfferedAmount
#2. OfferID accessed by O_Created, O_Sent, O_Returned O_Accepted O_Refused O_Cancelled O_Sent
#4. Everything else


cat(c("Rule 1: ", "not applicable", "\n"))
cat(c("Rule 2: ", unique(EL.OfferID.exists$Activity), "\n"))
cat(c("Rule 3: ", "FirstWithdrawalAmount, NumberOfTerms, Accepted, MonthlyCost, Selected, CreditScore, OfferedAmount", "\n"))
cat(c("Rule 4: ", "Variant, Variant.index, lifecycle.transition, EventOrigin, Action, LoanGoal, ApplicationType, RequestedAmount", "\n"))
```


### Community Structure Properties for Automated A2A Diagram Generation

Community Structure in graphs leads to clustering of connected vertices. We can take activities and attributes of an A2A Matrix as graph input.
This leads to clusters of related vertices (i.e., related attributes and activities), but identifying which of the vertices are attributes or activities might be non-trivial later. 

In the case of the offer application A2A Union Matrix, two clusters can be seen since the offer application event logs are representing two classes.
This version of graphing doesn't plot the graph as bipartite graph (i.e., the structure of an A2A diagram is unenforced).

```{r Graphing naive}
a2aMatrix <- a2a(Offer.union) # put any a2a matrix here to generate a2a diagram (Activity needs to be first column)
a2aMatrix <- column_to_rownames(a2aMatrix, 1) # make activities vertices
g <- graph_from_incidence_matrix(a2aMatrix, multiple = FALSE)

plot(g,
     vertex.label.cex=0.75, 
     vertex.label.family="Helvetica",
     vertex.label.font=1,
     
     vertex.shape="circle", 
     vertex.size=1, 
     vertex.label.color="black",
     edge.width=1,
     asp = 0.5,
     margin = 0)

```

By forcing the bipartite layout of A2A Diagrams, we can clearly identify related attributes.
In this case, the community structure of the A2A Diagram "finds" the same classes as MLAD + FCSE.



```{r Graphing bipartite}

l <- layout.bipartite(g, vgap = 0.1)
plot(g, layout=-l[,2:1],
     vertex.label.cex=0.75, 
     vertex.label.family="Helvetica",
     vertex.label.font=1,
     
     vertex.shape="circle", 
     vertex.size=1, 
     vertex.label.color="black",
     edge.width=1,
     asp = 0.5,
     margin = 0)
```


For the BPI 2020 logs this is less obvious, but still visible. The reduced BPI 2020 event logs have "correct" clustering.

Taking the Road Traffic Fine Management (RTFM) dataset from the original paper on discovering data models on single event logs, the community structure also "identifies" the same classes as the base algorithm.

```{r graphing RTFM}
RoadTraffic <- read.csv(here("data", "Road_Traffic_Fine_Management_Process.csv"))
RoadTraffic <- replaceEmptyWithNA(RoadTraffic)
RoadTraffic <- dropSuperfluousAttributes(RoadTraffic)

rt <- a2a(RoadTraffic) #put any a2a matrix here to generate a2a diagram (Activity needs to be first column)
rt <- column_to_rownames(rt, 1) #Remove Activity
g <- graph_from_incidence_matrix(rt, multiple = FALSE)
l <- layout.bipartite(g, vgap = 0.1)
plot(g, layout=-l[,2:1],
     vertex.label.cex=0.75, 
     vertex.label.family="Helvetica",
     vertex.label.font=1,
     
     vertex.shape="circle", 
     vertex.size=1, 
     vertex.label.color="black",
     edge.width=1,
     asp = 0.5,
     margin = 0)

```
